---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# imfr

<!-- badges: start -->
<!-- badges: end -->

Originally created by Christopher Gandrud, imfr is an R package for downloading data from the International Monetary Fund API endpoint. Version 2.0.0, by Christopher C. Smith, is an extensive revision of the package to make it both more powerful and more user-friendly. Regrettably, Version 2.0.0 is *not* backward-compatible; you will need to update old scripts if you update the package. The changes to the functionality of the package in this version were extensive, and adoption of the previous version was judged limited enough to make backward compatibility unnecessary.

## Installation

To install the development version of imfr, use:

``` r
devtools::install_github("chriscarrollsmith/imfr", build_vignettes = TRUE)
```
## Usage

I recommend using imfr in combination with the tidyverse, stringr, and knitr libraries, which introduces a powerful set of functions for viewing and manipulating the data types returned by imfr functions. Each of these packages can be installed from the CRAN repository using the `install.packages` function. Once they are installed, load these packages using the `library` function:

```{r load_libraries}
# Load libraries
library(imfr)
library(tidyverse)
library(stringr)
library(knitr)
```

The imfr package introduces four core functions: `imf_databases`, `imf_parameters`, `imf_parameter_defs`, and `imf_data`. The function for downloading datasets is `imf_data`, but you will need the other functions to determine what arguments to supply to `imf_data`. For instance, all calls to `imf_data` require a database_id. This is because the IMF serves many different databases through its API, and the API needs to know which of these many databases you're requesting data from. To obtain a list of databases, use `imf_databases`, like so:

```{r imf_databases}
#Fetch the list of databases available through the IMF API
databases <- imf_databases()
```

```{r count_invalid_dbs,echo=F,include=F}
library(tidyverse)
#Try to call each imf_database, and count the number that are invalid
parameter_defs <- map_dfr(databases$database_id,function(x){
    tryCatch({imf_parameter_defs(x)},error=function(cond){return(data.frame(parameter = NA, description = NA))}) %>%
        mutate(database = x)
}) 
invalid_dbs <- parameter_defs %>% filter(is.na(parameter))
```

This function returns the IMF's listing of `r nrow(databases)` databases available through the API. (In reality, `r nrow(invalid_dbs)` of the listed databases are defunct and not actually available: `r invalid_dbs$database`.)

To view and explore the database list, it's possible to open a viewing pane with `View(databases)` or to create an attractive table with `kintr::kable(databases)`. Or, if you already know which database you want, you can fetch the corresponding code by searching the description column for the database name with `stringr::str_detect`. For instance, here's how to search for the Primary Commodity Price System:

```{r commodity_id}
# Filter the 'databases' data frame for descriptions matching `commodity price`
commodity_db <- databases[str_detect(tolower(databases$description),"commodity price"),]

# Display the result using knitr::kable
kable(commodity_db)
```

Once you have a database_id, it's possible to make a call to `imf_data` to fetch the entire database: `imf_data(commodity_db$database_id)`. However, while this will succeed for some small databases, it will fail for many of the larger ones. And even when it succeeds, fetching an entire database can take a long time. You're much better off supplying additional filter parameters to reduce the size of your request.

Requests to databases available through the IMF API are complicated by the fact that each database uses a different set of parameters when making a request, and you also have to have the list of valid input codes for each parameter. You can obtain these using `imf_parameters`:

```{r imf_parameters}
# Fetch list of valid parameters and input codes for commodity price database
params <- imf_parameters(commodity_db$database_id)
```

The `imf_parameters` function returns a named list of data frames. Each named list item corresponds to a parameter used in making requests from the database. 

```{r explore_params}
# Check class of `params` object
class(params)

# Check names of `params` list items
names(params)
```

In the event that a parameter name is not self-explanatory, the `imf_parameter_defs` function can be used to fetch short text descriptions of each parameter:

```{r parameter_defs}
# Fetch and display parameter text descriptions for the commodity price database
param_descriptions <- imf_parameter_defs(commodity_db$database_id)
kable(param_descriptions)
```

Each named list item is a data frame containing a vector of valid input codes that can be used with the named parameter, and a vector of text descriptions of what each code represents. The `$` operator can be used to access the data frame for a given parameter, and the data frame can be explored using `kable` or `View`:

```{r freq_codes}
# Display the data frame of valid input codes for the frequency parameter
kable(params$freq)
```

There are two ways to supply parameters to `imf_data`: by supplying vector arguments or by supplying a modified parameters list.

To supply vector arguments, just find the codes you want and supply them to `imf_data` using the parameter name as the argument name. The example below shows how to request 2000–2015 annual coal prices from the Primary Commodity Price System database:

```{r annual_coal_vectors}
# Fetch the 'freq' input code for annual frequency
selected_freq <- params$freq$input_code[str_detect(tolower(params$freq$description),"annual")]

# Fetch the 'commodity' input code for coal
selected_commodity <- params$commodity$input_code[str_detect(tolower(params$commodity$description),"coal index")]

# Fetch the 'unit_measure' input code for index
selected_unit_measure <- params$unit_measure$input_code[str_detect(tolower(params$unit_measure$description),"index")]

# Request data from the API
df <- imf_data(database_id = commodity_db$database_id,
         freq = selected_freq, commodity = selected_commodity,
         unit_measure = selected_unit_measure,
         start_year = 2000, end_year = 2015)

# Display the first few entries in the retrieved data frame using knitr::kable
kable(head(df))
```

To supply a list object, modify each data frame in the `params` list object to retain only the rows you want, and then supply the modified list object to `imf_data` as its `parameters` argument. Here is how to make the same request for annual coal price data using a parameters list:

```{r annual_coal_list}
# Filter the frequency data frame for annual frequency
params$freq <- params$freq %>%
    filter(str_detect(tolower(.$description),"annual"))

# Filter the commodity data frame for the coal index
params$commodity <- params$commodity %>%
    filter(str_detect(tolower(.$description),"coal index"))

# Filter the unit_measure data frame for index
params$unit_measure <- params$unit_measure %>%
    filter(str_detect(tolower(.$description),"index"))

# Request data from the API
df <- imf_data(database_id = commodity_db$database_id,
               parameters = params,
         start_year = 2000, end_year = 2015)

# Display the first few entries in the retrieved data frame using knitr::kable
kable(head(df))
```

See also the vignettes, which can be accessed with `vignette("ParametersList")` and `vignette("ParametersVectors")`.